---
title: "Simple Model-Data Comparisons"
output: html_document
---

In this tutorial we will compare model outputs to data outside of the PEcAn web interface. The goal of this is to demonstrate that it is straightforward to perform additional analyses using PEcAn’s outputs. To do this you can download each of the Output files, and then perform the analyses using whatever software you prefer, or you can perform analyses directly on the PEcAn server itself. Here we’ll be analyzing model outputs in R using a browser-based version of RStudio that’s installed on the server

## Starting RStudio Server

1. Open RStudio Server in a new window at **URL/rstudio** 
 
2.	The username is carya and the password is illinois.

3.	To open a new R script click File > New File > R Script

4.	Use the Files browser on the lower right pane to find where your run(s) are located

  + All PEcAn outputs are stored in the output folder. Click on this to open it up.

  + Within the outputs folder, there will be one folder for each workflow execution. For example, click to open the folder PEcAn_99000000001 if that’s your workflow ID

  + A workflow folder will have a few log and settings files (e.g. pecan.xml) and the following subfolders

```  
run		contains all the inputs for each run
out		contains all the outputs from each run
pft		contains the parameter information for each PFT
```

Within both the run and out folders there will be one folder for each unique model run, where the folder name is the run ID. Click to open the out folder. For our simple case we only did one run so there should be only one folder (e.g. 99000000001). Click to open this folder.

  + Within this folder you will find, among other things, files of the format <year>.nc. Each of these files contains one year of model output in the standard PEcAn netCDF format.

## Load and plot the model output

```{r}
library(PEcAn.utils)
out <- read.output(99000000001, "output/PEcAn_99000000001/out/99000000001/" ,2004,2004,c("time","NEE"))
plot(out$time,out$NEE,type='l',ylab="kg m-2 s-1",xlab="day of year")
```

The arguments to read.output are the run ID, the folder where the run is located, the start year, the end year, and the variables being read.

##	Compare model to flux observations

**First** _load up the observations_ and take a look at the contents of the file

```{r}
library(PEcAn.assim.batch)
obs <- load.L2Ameriflux.cf("~/output/dbfiles/Ameriflux_site_0-622/US-Syv.2004.nc")
names(obs)
```

Here 0-622 is the database site ID, a folder that was created when you selected the site and PEcAn downloaded the Ameriflux NetCDF (.nc) file from the Ameriflux website. 

**Second**, _change all missing values_ from -9999 to not a number (NA) and also _$u*$ filter_ the fluxes. Here we will apply a conservative $u*$ of 0.2.

```{r}
obs[obs==-9999]=NA
obs$NEE[obs$UST<0.2]=NA
```

**Third**, _convert the model to have the same units as the data_

```{r}
modNEE <- ud.convert(out$NEE,"kg m-2 s-1","ug m-2 s-1")/12.0 #12 g C per mol
```

**Fourth**, _plot model predictions vs. observations_ and compare this to a 1:1 line 

```{r}
plot(modNEE,obs$NEE)
abline(0,1,col="red")  ## intercept=0, slope=1
```

**Fifth**, _calculate the Root Mean Square Error (RMSE)_ between the model and the data

```{r}
rmse = sqrt(mean((modNEE-obs$NEE)^2,na.rm = TRUE))
```

**Finally**, _plot time-series_ of both the model and data together

```{r}
plot(out$time,modNEE,type='l',ylab="umol m-2 s-1",xlab="day of year")
lines(out$time,obs$NEE,col="green")
```
