---
title: "ICOS Drought2018 Vignette"
author: 
- "Ayush Prasad"
- "Istem Fer"
date: "July 1, 2021"
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Drought-2018 ecosystem eddy covariance flux product

Drought-2018 ecosystem eddy covariance flux product is a release of the observational data product for eddy covariance fluxes at 52 stations from the Drought-2018 team, covering the period 1989-2018. 

For details see <https://www.icos-cp.eu/data-products/YVR0-4898>.

The following sites and years are available within this dataset:

```{r icos_d2018, echo=FALSE}
library(ggplot2)
icos_d2018 <- read.csv('extfiles/ICOS_Drought2018_Sites.csv')
knitr::kable(icos_d2018)
world_map <- ggplot2::map_data("world")
`%>%` <- dplyr::`%>%`
europe_cropped <-
        world_map %>% dplyr::filter((
          world_map$lat >= 35 & world_map$lat <= 75 &
            world_map$long >= -10 & world_map$long <= 40
        ))
# Compute the centroid as the mean longitude and lattitude
# Used as label coordinate for country's names
icos_data <- europe_cropped %>%
  dplyr::group_by(region) %>%
  dplyr::summarise(long = mean(long), lat = mean(lat))
ggplot(europe_cropped, aes(x = long, y = lat)) +
  geom_polygon(aes( group = group, color = region),fill="grey88")+
  scale_colour_manual(values=rep("black",length(unique(europe_cropped$region))),
                        guide=FALSE) +
  theme_void()+
  theme(legend.position = "none")+
  geom_point(data=icos_d2018, aes(x=Longitude, y=Latitude), color="red", size=2)

```

To download Drought2018 files we use `download.ICOS` function in PEcAn.
```{r download, eval=FALSE}
sitename   <- "FI-Hyy"
outfolder  <- "/home/carya/pecan"
start_date <- "2016-01-01"
end_date   <- "2018-01-01"
product <- "Drought2018"
res <- PEcAn.data.atmosphere::download.ICOS(sitename, outfolder, start_date, end_date, product)

```

To use the downloaded observations a function like `PEcAn.benchmark::load_data` can be used. Here we load the NEE observations.
```{r load, eval=FALSE}
 dbcon <-DBI::dbConnect(
    RPostgres::Postgres(),
    host = 'localhost',
    user = 'bety',
    password = 'bety',
    dbname = 'bety'
  )
 ## or if you have different DB settings try:
 ## php_config <- ".../pecan/web/config.php" # path to your PHP config file
 ## dcon <- betyConnect(php_config)
format <- PEcAn.DB::query.format.vars(bety = dbcon, format.id = 1000000136)
start_date <- lubridate::year("2018-01-01")
end_date   <- lubridate::year("2018-12-31")
vars.used.index <- which(format$vars$bety_name %in% c("NEE"))
obs <- PEcAn.benchmark::load_data(data.path = res$file, 
                                 format = format, start_year = start_date,  end_year = end_date,
                                  site = sitename, 
                                  vars.used.index = vars.used.index,
                                  time.row = format$time.row)

```

