---
title: "Create PostGIS polygon"
author: "David LeBauer"
date: "7/30/2015"
output: html_document
---

Code used to insert a polygon for Champaign County into a PostGIS geometry column.

Note that the `sites.geometry` column has three dimensions. Latitude and longitude are queried using `ggplot2::map_data('counties')` and elevation data is queried using `rgbif::elevation(latidutde, longitude)`.

This query is used to define the boundary of [site 1254, Champaign County](https://www.betydb.org/sites/1254), which will be used for testing and development of regional runs of PEcAn

### Load Libraries

```{r}
library(ggplot2)
library(data.table)
library(rgbif)
library(knitr)
```

### Select lat and lon points for Champaign County Boundary


```{r}
d <- data.table(map_data("county"))
champaign <- d[region == "illinois" & subregion == "champaign"]

kable(champaign)
```

### Query Elevation points

```{r}
### for PostGIS, need polygon to return to start point
champaign <- rbind(champaign, champaign[1])

## need to get key from Google Maps API
## I've put mine in a ~/.googlemapskey
key <- readLines(con = "~/.googlemapskey")
champaign_pts <- data.table(champaign[,elevation(latitude = lat, longitude = long, key = key)])

kable(champaign_pts, caption = "Table with elevations")
```


### Define boundary in postGIS

```{r}
boundary <- champaign_pts[,paste(longitude, latitude, elevation, collapse = ",")]

writeLines(boundary)
```

### Define PostGIS geom

Use `srid = 4326`

```{r}
geometry <- paste0("ST_SetSRID(ST_MakePolygon(ST_GeomFromText('LINESTRING(", boundary , ")')), 4326)")

writeLines(geometry)
```

#### insert statement

```{r}
insert_polygon <- paste0("INSERT into sites (sitename, geometry) VALUES ('Champaign County', ", geometry, ");")
writeLines(insert_polygon)
```

#### Update geom statement

```{r}
update_polygon <- paste0("update sites set geometry = ", geometry , " where sitename = 'Champaign County';")
writeLines(update_polygon)
```


### Insert into BETYdb

```{r, eval = FALSE}

library(PEcAn.DB)
library(RPostgreSQL)
#settings <- read.settings('settings.xml')
settings <-list(database = list(bety = list(driver = "PostgreSQL", user = "bety", dbname = "bety", password = "bety")))

db.query(insert_polygon, con = db.open(settings$database$bety))
```