---
title: "Benchmarking Demo"
author: "Betsy Cowdery"
date: "August 9, 2016"
output: html_document
runtime: shiny
---
## Setup 

```{r,message=FALSE,warning=FALSE}

options(digits = 10) # just to make things easier to read

library(PEcAn.all)
library(PEcAn.benchmark)
library(RPostgreSQL)
library(XML)
library(dplyr)

#db connection and host information

dbparms <- list(
  user = "bety",
  password = "bety",
  host = "psql-pecan.bu.edu",
  dbname = "bety",
  write = TRUE,
  driver = "PostgreSQL"
)
con     <- db.open(dbparms)
user_id <- 1000000003 # Betsy's ID 

# In case you want to re-run calc.benchmark for an existing benchmark ensemble run - but if that's the case then you should probably start the workflow with the given bm.enseble id. This may just end up being for testing. 
new.run <-  FALSE

# Example ensemble.ids to try
ensemble.id <- 1000003655
```

## Definition
### By hand: Formats, Inputs and Database files

There are parts of the process that are not fully automated yet so before proceeding, make sure you do these steps:

1) Create records for the benchmarking data sources you are going to use in the formats, inputs and dbfiles tables. 

For example, my format record looks like this:
```{r}
print(db.query("SELECT * FROM formats WHERE name = 'NPP_Synthesis_CDIAC'",con))
```
then my input record looks like this:
```{r}
print(db.query("SELECT * FROM inputs WHERE id = 1000000651",con))
```
and finally my database file record looks like this:
```{r}
print(db.query("SELECT * FROM dbfiles WHERE container_id = 1000000651 AND container_type = 'Input'",con))
```

2) Create new benchmark etries for variables of interest in the new data sources.

For example, I made two benchmark entries that look like this:
```{r}
print(db.query("SELECT * FROM benchmarks WHERE input_id = 1000000651",con))
```

### Benchmark Reference Run and Ensemble

```{r}
# get information about the ensemble and workflow 
ensemble <- db.query(paste("SELECT * FROM ensembles where id = ",ensemble.id,";"), con)
workflow <- db.query(paste("SELECT * FROM workflows where id = ",ensemble$workflow_id,";"), con)

# check if there is already a BRR for ensemble.id
BRR <- db.query(sprintf("SELECT * FROM reference_runs where settings = '%s' ", workflow$folder), con)
if(length(BRR) == 0){
  BRR <- create.BRR(ensemble.id, workflow, con)
} 

bm.ensembles <- db.query(sprintf("SELECT * FROM benchmarks_ensembles WHERE reference_run_id = %0.f ", BRR$id),con)
```

If you are making a new benchmark reference run, then you will have to relate you existing benchmarks with the reference run in the benchmarks_benchmarks_reference_runs table.

```{r}
# benchmarks_BRR <- db.query(paste(
#   "INSERT INTO benchmarks_benchmarks_reference_runs (benchmark_id, reference_run_id)",
#   "VALUES (",benchmarks$id,",",BRR$id,") RETURNING * "),con)

print(db.query(paste("SELECT * FROM benchmarks_benchmarks_reference_runs WHERE reference_run_id = ",BRR$id),con))

bm.ensembles <- db.query(sprintf("SELECT * FROM benchmarks_ensembles WHERE reference_run_id = %0.f ", BRR$id),con)
```

### New Runs (currently not functional)
```{r}

bm.host <- db.query(sprintf("SELECT w.hostname, e.id FROM workflows as w JOIN ensembles as e ON w.id = e.workflow_id JOIN benchmarks_ensembles as be ON e.id = be.ensemble_id WHERE be.reference_run_id = %0.f ", BRR$id),con)
local <- which(bm.host$hostname == fqdn())

if(length(local) == 0 | new.run == TRUE){
  bm.ensemble <- start.bm.ensemble(BRR, con)
}else{
  # how do you choose which ref run to use? Maybe through shiny? For now, take the 1st one...
  bm.ensemble <- filter(bm.ensembles, ensemble_id == bm.host$id[local][1])
}
```
## Evaluation
### Calculations

```{r,echo=FALSE}

results <- calc.benchmark(bm.ensemble, con)

```

### Reports

Ultimately should launch a nicer shiny app
print(source('modules/benchmark/R/prototype.shiny.2.R')$value)

```{r}

dat.vars <- setdiff(names(results$input.1000000651$bench.results), "metric")
metrics <- results$input.1000000651$bench.results$metric
metric.plot <- metrics[grep("plot",metrics)]

ui <- shinyUI(pageWithSidebar(
  headerPanel = ("Benchmarking Visualization Prototype"),
  sidebarPanel(
    conditionalPanel(condition="input.conditionedPanels==1",
                     selectInput(inputId = "b", label = "Benchmark Data Set", choices = names(results)),
                     selectInput(inputId = "v", label = "Variable", choices = dat.vars),
                     radioButtons(inputId = "metric.plot", label="What plot you like to see?", 
                                  choices = metric.plot)
    ),
    conditionalPanel(condition="input.conditionedPanels==2",
                     selectInput(inputId = "b", label = "Benchmark Data Set", choices = names(results))
    ),
    conditionalPanel(condition="input.conditionedPanels==3",
                     selectInput(inputId = "b", label = "Benchmark Data Set", choices = names(results)),
                     selectInput(inputId = "v", label = "Variable", choices = dat.vars)
    )
  ),
  mainPanel(
    tabsetPanel(
      tabPanel("Plot", plotOutput("p"), value = 1), 
      tabPanel("Scores", dataTableOutput('scores'), value=2), 
      tabPanel("Benchmarking Inputs", dataTableOutput('obvs'), value=2),
      tabPanel("Model Outputs", dataTableOutput('model'), value=2),
      tabPanel("Aligned data", dataTableOutput('dat'), value=3),
      tabPanel("Formats", dataTableOutput('format'), value=2), 
      id = "conditionedPanels"
    )
    
  )
)
)


server <- function(input, output){
  output$p <- renderPlot({
    metric_dat <- results[[input$b]]$aligned.dat[[input$v]]
    colnames(metric_dat)<- c("model","obvs","time")
    fcn <- paste0("metric.",input$metric.plot)
    do.call(fcn, args <- list(metric_dat,input$v,draw.plot=TRUE))
  })
  output$scores <- renderDataTable(results[[input$b]]$bench.results[-grep("plot",metrics),])
  output$obvs  <- renderDataTable(results[[input$b]]$obvs[,-which(colnames(results[[input$b]]$obvs)=="posix")])
  output$model <- renderDataTable(as.data.frame(results[[input$b]]$model[,-which(colnames(results[[input$b]]$model)=="posix")]))
  output$dat <- renderDataTable(as.data.frame(results[[input$b]]$aligned.dat[[input$v]][,-which(colnames(results[[input$b]]$aligned.dat[[input$v]])=="posix")]))
  output$format <- renderDataTable(results[[input$b]]$format)
}

shinyApp(ui=ui, server=server, options = list(width = "100%", height = 800))

```

