---
title: "Site Selection"
author: "MCD"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(terra)
```

This document contains the code for the initial prep of a landscape-scale run

* Settings: bounding box and grid resolution
* Inputs: data layers used for assigning locations to a "patch"
* Outputs: grid of patch assignment IDs used to process inputs and data constraints

## Settings
```{r}
grain = 30
tower_latlon = c(42.53691, -72.17265)
tower_lonlat = rev(tower_latlon)
tower_utm    = c(732183.98,4713266.18)
nclass = 10
```

# Inputs


## NEON LAI
NEON AOP mosaic data are provided in tiles with UTM coordinates in each file name. This is not explained well on the NEON data portal

NEON (National Ecological Observatory Network). LAI - spectrometer - mosaic (DP3.30012.001), RELEASE-2022. https://doi.org/10.48443/ys43-cb22. Dataset accessed from https://data.neonscience.org on December 21, 2022

```{r}
## load data
laifile = "NEON_lai-spectrometer-mosaic/NEON.D01.HARV.DP3.30012.001.2019-08.basic.20221221T142516Z.RELEASE-2022/NEON_D01_HARV_DP3_732000_4713000_LAI.tif"
lairaw <- rast(laifile)
laie <- rast("NEON_lai-spectrometer-mosaic/NEON.D01.HARV.DP3.30012.001.2019-08.basic.20221221T142516Z.RELEASE-2022/NEON_D01_HARV_DP3_732000_4713000_LAI_Error.tif")
```

First download only got one tile, went back for more!
NEON (National Ecological Observatory Network). LAI - spectrometer - mosaic (DP3.30012.001), RELEASE-2023. https://doi.org/10.48443/7x5a-mn68. Dataset accessed from https://data.neonscience.org on March 22, 2023

Need to mosaic together: https://www.neonscience.org/resources/learning-hub/tutorials/merge-aop-raster-data
using code https://raw.githubusercontent.com/NEONScience/NEON-Data-Skills/main/tutorials/R/AOP/AOP-L3-rasters/aop_merge_raster_functions.R
```{r}
source("~/pecan/modules/assim.sequential/inst/hf_landscape/aop_merge_raster_functions.R")
lai_folder = "/projectnb/dietzelab/dietze/hf_landscape_SDA/NEON_lai-spectrometer-mosaic/NEON.D01.HARV.DP3.30012.001.2019-08.basic.20230322T174312Z.RELEASE-2023/"

#get the list of dataTiles for each data directory and extension
dataTiles <- getDataTilesByExt(lai_folder,dataExts = getDataExts("DP3.30012.001"))

cat('Creating full site mosaics\n')
fullMosaics <- list()
fullMosaicNames <- list()
for (i in 1:length(dataTiles)) {
  fullMosaics[[i]] <- mergeDataTiles(dataTiles[[i]])
  dataDirSplit <- unlist(strsplit(dataTiles[[i]][1],.Platform$file.sep))
  tileNameSplit <- unlist(strsplit(dataTiles[[i]][1],'_'))
  if (tail(tileNameSplit,1)=='error.tif') {
    fullMosaicNames[[i]] <- paste0(dataDirSplit[9],'_',tail(tileNameSplit,n=2)[1],'_error.tif')} 
  else {
    fullMosaicNames[[i]] <- paste0(dataDirSplit[9],'_',tail(tileNameSplit,n=1))}
}

cat('Writing full mosaics to Geotiffs and Cloud-Optimized Geotiffs\n')
outFileDir = file.path(dirname(lai_folder),"mosaic")
makeDir(outFileDir)
for (i in 1:length(fullMosaics)) {
  outFileTif <- fullMosaicNames[[i]]
  cat(paste0('Generating ',outFileTif,'\n'))
  writeFullMosaicTif(fullMosaics[[i]],outFileDir,outFileTif)
  # name the COG file the same as tif but with COG suffix
  outFileCog <- gsub(".tif", "_COG.tif", outFileTif) 
  cat(paste0('Generating ',outFileCog,'\n'))
  convertTif2Cog(outFileDir,outFileTif,outFileCog)
}
dir(outFileDir)

## load file
laifile = data.frame(file=dir(outFileDir,full.names = TRUE)) %>% dplyr::filter(!(grepl("Error",file) | grepl("COG",file)))
lairaw <- rast(laifile$file)
```


## regrid
```{r}
#lai <- aggregate(lairaw,fact=grain,na.rm=TRUE)
```



```{r}
## visualize data
lairaw2 = terra::clamp(lairaw,upper=5)
plot(lairaw2)
points(tower_utm[1],tower_utm[2])

#plot(lai)
#points(tower_utm[1],tower_utm[2])
```
Set overall grid  based on LAI map
```{r}
box = ext(lai)
```

## NLCD
```{r}
url = "https://cpdataeuwest.blob.core.windows.net/cpdata/raw/nlcd/conus/30m/2016.tif"  ## check with Carl about why this didn't work
url = "/projectnb/dietzelab/dietze/nlcd/2016.tif"
nlcd_full = rast(url)
```

```{r}
## move NLCD and LAI to same grain and extent
lairp = project(lairaw2,crs(nlcd_full))
nlcd = crop(nlcd_full,lairp)
lai <- exactextractr::exact_resample(lairp, nlcd, 'mean')

plot(nlcd)
plot(lai)
plot(nlcd,lai)

```

## landtrendr agb
```{r}
agbpath = "/projectnb/dietzelab/dongchen/Multi-site/download_500_sites/AGB/biomassfiaald_2017_median.tif"
agb_full <- rast(agbpath)
agb = crop(agb_full,nlcd)
agb = project(agb,crs(nlcd))

plot(agb,lai)

```

## site selection 1: NLCD
```{r}
## remove non-vegetated pixels
nlcd[nlcd < 41] = NA
nlcd[is.na(agb)] = NA
nlcd[is.na(lai)] = NA
plot(nlcd)

## repeat until nothing is rare or things cant be merged
n_nlcd = table(as.vector(nlcd))
n_nlcd = n_nlcd/sum(n_nlcd)

## aggregate rare classes
rare = names(n_nlcd)[which(n_nlcd*nclass < 0.5)]

last = names(which.min(n_nlcd))
if(last %in% c(72:82,95)) nlcd[nlcd == as.numeric(last)] = 71 ## merge herbaceous
if(last %in% c(52)) nlcd[nlcd == as.numeric(last)] = 43       ## merge scrub

writeRaster(nlcd, "nlcd_crop.tif", overwrite=TRUE)
writeRaster(lai, "lai_crop.tif", overwrite=TRUE)
writeRaster(agb, "agb_crop.tif", overwrite=TRUE)

```

## determine splits of common classes
```{r}
n_nlcd = table(as.vector(nlcd))
n_nlcd = n_nlcd/sum(n_nlcd)
grouped = names(n_nlcd)[which(n_nlcd < 1/nclass)]

## repeat until all assigned
common = n_nlcd[!(names(n_nlcd) %in% grouped)]
last = names(which.min(common))
remaining_area = sum(common)
remaining_bins = nclass - length(grouped)
nlast = round(common[which(names(common) == last)] /  (remaining_area/remaining_bins))
if(length(common)==1) nlast = remaining_bins
grouped = c(grouped, rep(last,nlast))

### sanity checks
length(grouped) == nclass

table(grouped)/nclass
n_nlcd

```
## cluster and split the common classes
```{r}
nlcd[is.na(nlcd)] = 0

reps = table(grouped)
common = as.numeric(names(reps)[reps > 1])

for(i in seq_along(common)){
  nlcdc = as.vector(nlcd)[which(as.vector(nlcd) == common[i])]
  laic = scale(as.vector(lai)[which(as.vector(nlcd) == common[i])])
  agbc = scale(as.vector(agb)[which(as.vector(nlcd) == common[i])])
  df = cbind(laic,agbc)
  #sel = apply(!is.na(df),1,all)
  
  k = kmeans(df, centers = reps[names(reps) == common[i]], nstart = 25)
  plot(df,col=k$cluster)
  
#  nlcd[nlcd == common[i]] = common[i] + k$cluster/nclass
  nlcd[abs(nlcd - common[i]) < 0.3] = common[i] + k$cluster/nclass
}
#nlcd = as.numeric(nlcd)
table(as.vector(nlcd))

nlcd[nlcd == 0] = NA
```

## recode and generate summary stats
```{r}
area   = table(as.vector(nlcd))
area   = area/sum(area)
common = names(area)
mylai = tapply(as.vector(lai),as.vector(nlcd),mean,na.rm=TRUE)
myagb   = tapply(as.vector(agb),as.vector(nlcd),mean,na.rm=TRUE)

df = tibble::tibble(patch = seq_along(common),
            area  = area,
            nlcd  = floor(as.numeric(common)),
            lai = mylai,
            agb = myagb
           )
write.csv(df,file = "patches.csv",row.names = FALSE,col.names = TRUE)

size = df$area*nclass*1.5
size[size<0.5] = 0.5
plot(df$agb,df$lai,type="n")
text(df$agb,df$lai,as.character(df$nlcd),cex=size)

```


```{r}
for(i in seq_along(common)){
  nlcd[nlcd == as.numeric(common[i])] = i
}
writeRaster(nlcd, "patches.tif", overwrite=TRUE)
```


## compare patches to ground plots
```{r}
library(dplyr)
patches = df
if(!exists("pmap")){
  patchMap <- terra::rast("patches.tif")
  pmap <- project(patchMap, "+proj=longlat",method="near")
  plot(pmap)
  points(tower_lonlat[1],tower_lonlat[2])
  tow = extract(pmap,tower_lonlat)
}

pfile = "/projectnb/dietzelab/dietze/hf_landscape_SDA/HF_Biometry_Plots_Lat_Long.csv"
if(file.exists(pfile)){
  inv.plot = read.csv(pfile) %>% rename(lat=Latitude,lon=Longitude) %>% select(lon,lat)
  inv.patch = extract(pmap,inv.plot)  
  count = table(inv.patch$Layer_1)
  count = tibble(patch=as.numeric(names(count)),n=count)
  p2 = left_join(patches,count,"patch")
  
  plot(pmap)
  points(inv.plot$lon,inv.plot$lat)
}

lairaw3 = terra::project(lairaw2,"+proj=longlat")
plot(lairaw3)
points(inv.plot$lon,inv.plot$lat)

inv.ll = terra::vect(inv.plot, crs="+proj=longlat")
plot(lairaw3)
points(inv.ll)

inv.utm = terra::project(inv.ll,lairaw2)
plot(lairaw2)
points(inv.utm)
```


