---
title: "PEcAn Validation Framework: Comprehensive User Guide & Tutorial"
author: "PEcAn Project"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PEcAn Validation Framework: Comprehensive User Guide & Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4,
  warning = FALSE,
  message = FALSE
)
```

# Introduction

The **PEcAn General-Purpose Validation Toolkit** (`PEcAn.benchmark`) provides a database-decoupled, modular library of functions for comparing ecological model predictions against observed site-level and regional datasets.

Unlike legacy benchmarking workflows tied directly to BETYdb database schemas, the modern validation toolkit operates natively on standard R `data.frame` structures. This enables researchers to rapidly ingest offline CSV or NetCDF datasets, perform spatiotemporal alignment, evaluate ensemble uncertainty metrics, and compile interactive ILAMB-style HTML validation scorecards.

This vignette provides a step-by-step tutorial covering:

1. Expected data schemas (variables, units, and ensemble representations).
2. Ground-truth simulated observations generated from a known statistical error model.
3. Data preparation, temporal aggregation, and time alignment.
4. Validation functions walkthrough (ensemble mean vs. full ensemble UQ metrics).
5. Visualizations, scorecards, and report generation.
6. User data substitution guide for custom datasets.

---

# 1. Inputs and Expected Data Schemas

The validation framework expects two primary data structures in memory or stored as plain text/NetCDF files:

### Observation Schema
* **`time`**: Datetime object (`POSIXct` or ISO-8601 formatted timestamp).
* **`obvs`**: Numeric vector of observed values for target variable $v$.
* **`obvs_sd`**: (Optional) Standard error or measurement uncertainty of observation.
* **`variable`**: Character label for variable (e.g., `"NEE"`, `"GPP"`, `"SOC"`).
* **`unit`**: Physical measurement unit (e.g., `"umol m-2 s-1"`, `"kg C m-2"`).

### Model Output Schema
The framework supports both **single model runs** and **ensemble model outputs**:
* **Single Model Run**: `time`, `model` (single numeric vector of predictions).
* **Ensemble Model Run**: 
  * Individual member columns: `model_1`, `model_2`, ..., `model_N`.
  * Summary statistics: `ensemble_mean` ($m_{v,t}$), `ensemble_sd`, `model_q2.5` (2.5% quantile), and `model_q97.5` (97.5% quantile).

---

# 2. Simulated Observations with Known Error Model

To demonstrate the framework and verify metric calculations against theoretical ground truth, we generate synthetic observations $o_{v,t}$ from ensemble mean outputs $m_{v,t}$ using a known statistical error model:

$$o_{v,t} = m_{v,t} + S_v (\beta_v + \epsilon_{v,t}), \quad \epsilon_{v,t} \sim N(0, \tau_v^2)$$

Where:
* $o_{v,t}$ is the simulated observation for variable $v$ at timestep $t$.
* $m_{v,t}$ is the ensemble mean prediction after aggregation.
* $S_v$ scales the error relative to the variable's variability ($S_v = \text{sd}(m_{v,t})$).
* $\beta_v$ is the relative bias parameter ($\beta_v = 0.30$, i.e., 30% of standard deviation).
* $\tau_v$ controls random error ($\tau_v = 0.50$, i.e., 50% of standard deviation).

### Side Note: Code Used to Generate Precomputed Fixtures

Observations and model ensemble predictions are precomputed and stored as plain-text inspectable CSV files (`inst/extdata/simulated_observations.csv` and `inst/extdata/simulated_model_ensemble.csv`).

```r
# Side note: Statistical observation generator code
set.seed(3)

output_scale <- stats::sd(ensemble_mean) # S_v

# Parameterization: Relative bias = 30% of SD, Relative error = 50% of SD
relative_bias  <- 0.3
relative_error <- 0.5

simulated_observation <- ensemble_mean +
  output_scale * (
    relative_bias +
      stats::rnorm(
        length(ensemble_mean),
        mean = 0,
        sd = relative_error
      )
  )
```

### Loading Precomputed Text Fixtures

```{r load_fixtures}
if (!suppressWarnings(requireNamespace("PEcAn.benchmark", quietly = TRUE))) {
  devtools::load_all(file.path(".."), quiet = TRUE)
} else {
  library(PEcAn.benchmark)
}

# Path to precomputed plain-text CSV fixtures
extdata_path <- system.file("extdata", package = "PEcAn.benchmark")

if (extdata_path == "" || !file.exists(file.path(extdata_path, "simulated_observations.csv"))) {
  # Fallback to relative directory if package is loaded via devtools/rmarkdown render
  extdata_path <- file.path("..", "inst", "extdata")
}

obs_path <- file.path(extdata_path, "simulated_observations.csv")
model_path <- file.path(extdata_path, "simulated_model_ensemble.csv")

# Read plain text fixtures
obs_df <- read.csv(obs_path)
model_df <- read.csv(model_path)

# Convert timestamps to POSIXct
obs_df$time <- as.POSIXct(obs_df$time, tz = "UTC")
model_df$time <- as.POSIXct(model_df$time, tz = "UTC")

head(obs_df, 3)
head(model_df[, c("time", "ensemble_mean", "model_q2.5", "model_q97.5")], 3)
```

---

# 3. Preparation: Temporal Aggregation & Alignment

Ecosystem models and tower observation feeds often record timestamps with minor temporal offsets or different sample frequencies.

The `align_by_time()` function aligns model and observation data frames using binary search (`findInterval`) within a user-specified tolerance window (in seconds).

```r
# Align model and observation data within a 1-hour window (3600 seconds)
aligned_df <- align_by_time(
  model_df = model_df,
  obs_df = obs_df,
  tolerance_secs = 3600
)

# Inspect aligned dataframe
head(aligned_df[, c("time", "obvs", "ensemble_mean", "model_q2.5", "model_q97.5")])
```

---

# 4. Validation Functions & Metric Walkthrough

The framework segregates validation metrics into two distinct categories based on whether they assess point predictions or uncertainty calibration:

### Category A: Ensemble Mean (Point Prediction) Metrics
These metrics compare the observation $o_{v,t}$ against the ensemble mean $m_{v,t}$:

1. **Bias (`metric_bias`)**: Systematics shift / offset.
   $$\text{Bias} = \frac{1}{N}\sum (m_{v,t} - o_{v,t})$$
   *Expected Theoretical Value:* $\approx -S_v \cdot \beta_v$.

2. **Root Mean Squared Error (`metric_RMSE`)**: Overall magnitude of prediction error.
   $$\text{RMSE} = \sqrt{\frac{1}{N}\sum (m_{v,t} - o_{v,t})^2}$$

3. **Mean Absolute Error (`metric_MAE`)**: Average absolute deviation (less sensitive to extreme outliers than RMSE).
   $$\text{MAE} = \frac{1}{N}\sum |m_{v,t} - o_{v,t}|$$

4. **Coefficient of Determination ($R^2$) (`metric_R2`)**: Proportion of variance explained.

5. **Nash-Sutcliffe Model Efficiency (NSE/MEF) (`metric_NSE`)**: Model skill relative to mean observation benchmark ($NSE = 1 \implies \text{perfect match}$, $NSE < 0 \implies \text{mean observation is better predictor}$).

```r
# Compute Point / Ensemble Mean Metrics
bias_val <- metric_bias(aligned_df)
rmse_val <- metric_RMSE(aligned_df)
mae_val  <- metric_MAE(aligned_df)
r2_val   <- metric_R2(aligned_df)

cat(sprintf("Bias:  %.4f\n", bias_val))
cat(sprintf("RMSE:  %.4f\n", rmse_val))
cat(sprintf("MAE:   %.4f\n", mae_val))
cat(sprintf("R^2:   %.4f\n", r2_val))
```

---

### Category B: Full Ensemble (Uncertainty Quantification) Metrics
These metrics evaluate whether the ensemble spread accurately reflects predictive uncertainty:

1. **Prediction Interval Coverage (`metric_Coverage`)**:
   Calculates the empirical percentage of observations that fall within the model's nominal 95% predictive interval (`model_q2.5` to `model_q97.5`).
   *Interpretation:* A well-calibrated ensemble yields $\approx 95\%$ coverage. Coverage $< 95\%$ indicates underconfident/underdispersed uncertainty bounds.

2. **Predictive Model Uncertainty (PMU) (`metric_PMU`)**:
   Ratio of ensemble variance to total predictive variance.

3. **Continuous Ranked Probability Score (`metric_CRPS`)**:
   Evaluates the full predictive probability distribution against observed scalar outcomes. Generalizes MAE to probabilistic forecasts.

```r
# Compute Full Ensemble UQ Metrics
coverage_val <- metric_Coverage(aligned_df)
pmu_val      <- metric_PMU(aligned_df)
crps_val     <- metric_CRPS(aligned_df)

cat(sprintf("95%% Interval Coverage: %.2f%%\n", coverage_val * 100))
cat(sprintf("PMU Score:             %.4f\n", pmu_val))
cat(sprintf("Ensemble CRPS:         %.4f\n", crps_val))
```

---

# 5. Visualizations, Scorecards & Validation Report

The framework provides visualization wrappers built on `ggplot2` to evaluate temporal trends, residuals, and correlation.

### Timeseries Plot with Ensemble Uncertainty Ribbons
```r
ts_plot <- metric_timeseries_plot(
  metric_dat = aligned_df,
  var = "Net Ecosystem Exchange (NEE)",
  unit = "umol m-2 s-1",
  draw.plot = TRUE
)
```

### Residual Analysis Plot
```r
res_plot <- metric_residual_plot(
  metric_dat = aligned_df,
  var = "NEE Residuals",
  unit = "umol m-2 s-1",
  draw.plot = TRUE
)
```

### Modeled vs. Observed Scatter Plot
```r
scatter_plot <- metric_scatter_plot(
  metric_dat = aligned_df,
  var = "Modeled vs Observed NEE",
  unit = "umol m-2 s-1",
  draw.plot = TRUE
)
```

### Automated Quarto Scorecard Generation
To generate a self-contained HTML report (similar to ILAMB scorecards):

```r
# Package benchmark results
benchmark_results <- list(
  metrics = data.frame(
    Site = "Simulated_Site_01",
    Variable = "NEE",
    Bias = bias_val,
    RMSE = rmse_val,
    MAE = mae_val,
    R2 = r2_val,
    Coverage_Pct = coverage_val * 100,
    CRPS = crps_val
  ),
  aligned_data = aligned_df,
  plots = list(
    timeseries = ts_plot,
    residuals = res_plot,
    scatter = scatter_plot
  )
)

# Render automated HTML scorecard report
generate_validation_report(
  benchmark_results = benchmark_results,
  output_file = "Simulated_Validation_Report.html"
)
```

---

# 6. User Data Substitution Guide

Users can substitute their own processed observations and model outputs into the validation framework using standard file loaders or custom data frame mapping:

### Option A: Loading Custom CSV Files
```r
# Map custom CSV headers to PEcAn framework vocabulary
custom_obs <- load_csv(
  path = "my_tower_data.csv",
  format = list(
    time = "TIMESTAMP",
    obvs = "FC_F_MDS",
    obvs_sd = "FC_F_MDS_QC"
  )
)
```

### Option B: Loading NetCDF Model Outputs
```r
# Load NetCDF model predictions
custom_model <- load_x_netcdf(
  path = "my_model_output.nc",
  vars = c("NEE", "GPP")
)
```

### Option C: Direct In-Memory DataFrame Substitution
If your data is already in memory:

```r
my_aligned_data <- data.frame(
  time = my_timestamps,
  obvs = my_observations,
  ensemble_mean = my_model_means,
  model_q2.5 = my_lower_quantiles,
  model_q97.5 = my_upper_quantiles
)

# Pass directly to validation functions
my_rmse <- metric_RMSE(my_aligned_data)
my_coverage <- metric_Coverage(my_aligned_data)
```

---

# 7. Multi-Site Ensemble Fixture Example (Salinas SOCs)

The package includes a lightweight multi-site ensemble test fixture (`small_salinas_ensemble.csv` and `small_salinas_obs.csv`) containing predictions across 2 management systems (`socs_sys1`, `socs_sys2`), 2 years of observations (2010–2011), and 10 ensemble members per timestep.

This fixture demonstrates how the framework handles multi-site ensemble matrices, temporal alignment, and multi-site scorecard computation:

```{r small_fixture_example}
# Load small multi-site Salinas SOCs fixture (2 systems, 2 years, 10 ensemble members)
if (!exists("extdata_path") || !file.exists(file.path(extdata_path, "small_salinas_ensemble.csv"))) {
  extdata_path <- system.file("extdata", package = "PEcAn.benchmark")
  if (extdata_path == "" || !file.exists(file.path(extdata_path, "small_salinas_ensemble.csv"))) {
    extdata_path <- file.path("..", "inst", "extdata")
  }
}

small_ens_path <- file.path(extdata_path, "small_salinas_ensemble.csv")
small_obs_path <- file.path(extdata_path, "small_salinas_obs.csv")

if (file.exists(small_ens_path) && file.exists(small_obs_path)) {
  ens_data <- read.csv(small_ens_path)
  obs_data <- read.csv(small_obs_path)

  # Reshape ensemble data to wide format per site
  sites <- unique(ens_data$site_id)
  site_scorecards <- list()

  for (s in sites) {
    site_ens <- ens_data[ens_data$site_id == s, ]
    site_obs <- obs_data[obs_data$site_id == s, ]

    # Pivot ensemble members to matrix via PEcAn.benchmark utility
    ens_mat <- efi_long_to_array(site_ens)
    time_seq <- unique(site_ens$datetime)

    m_summary <- data.frame(
      time = as.POSIXct(time_seq, tz = "UTC"),
      value = rowMeans(ens_mat),
      model_q05 = apply(ens_mat, 1, quantile, probs = 0.05),
      model_q95 = apply(ens_mat, 1, quantile, probs = 0.95),
      site = s
    )

    o_summary <- data.frame(
      time = as.POSIXct(site_obs$date, tz = "UTC"),
      value = site_obs$obs_mean,
      obvs_sd = site_obs$obs_sd,
      site = s
    )

    aligned <- align_by_time(m_summary, o_summary, tolerance_secs = 365 * 86400 / 2)
    aligned_indices <- match(aligned$time, m_summary$time)
    attr(aligned, "ensemble_matrix") <- ens_mat[aligned_indices, , drop = FALSE]

    site_scorecards[[s]] <- compute_metrics(aligned, metrics = c("RMSE", "MAE", "R2", "COVERAGE", "CRPS"))
  }

  multi_site_scorecard <- do.call(rbind, site_scorecards)
  print(multi_site_scorecard)
}
```

---

# Conclusion

The PEcAn General-Purpose Validation Toolkit offers an adaptable, database-decoupled framework for rigorous ecological model benchmarking. By separating point prediction metrics (ensemble mean) from distribution metrics (full ensemble UQ) and providing automated Quarto scorecards, researchers can evaluate model performance across sites and ecosystem scales.
