---
title: "PEcAn Benchmarking and Validation Framework Tutorial"
author: "PEcAn Development Team"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PEcAn Benchmarking and Validation Framework Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Introduction

The `PEcAn.benchmark` toolkit provides a modular framework for aligning, evaluating, and visualizing ecosystem model predictions against observational benchmark datasets.

This tutorial walks through the complete workflow of the validation framework:
1. Understanding data intake schemas and ensemble representation.
2. Generating simulated observations from a statistical error model with known parameters.
3. Reshaping EFI long model predictions and temporally aligning model and observation data.
4. Computing statistical metrics on both the ensemble mean (point metrics) and the full ensemble spread (probabilistic metrics).
5. Generating diagnostic plots (ribbon/spaghetti, scatter, residual plots) and rendering an automated Quarto validation scorecard.
6. Substituting custom model predictions and observational data.

---

## 1. Input Data Schemas and Ensemble Representation

The validation toolkit expects model outputs and observations in standardized formats.

### Model Ensemble Output (EFI Long Format)
Model ensemble outputs follow the EFI (Ecological Forecasting Initiative) standard long format:

* `datetime`: Timestamp string (e.g. `"2020-01-01 00:00:00"`).
* `site_id`: Site identifier (e.g. `"socs_sys1"` or `"US-Ha1"`).
* `parameter`: Integer index or member ID for each ensemble member (e.g. `1, 2, ..., N`).
* `variable`: CF/PEcAn standard variable name (e.g. `"TotSoilCarb"`, `"NEE"`, `"GPP"`).
* `prediction`: Model output numeric prediction value.

### Observation Dataset
Observational data require at minimum:
* `date` / `time`: Timestamp (POSIXct or parseable string).
* `obs_mean` / `obvs`: Observational numeric mean value.
* `obs_sd` / `obvs_sd` (optional): Observational standard deviation / measurement error.
* `site_id` / `site` (optional): Site identifier matching the model dataset.

---

## 2. Statistical Error Model and Fixture Generation

To evaluate model metrics against known true error parameterizations, simulated observations can be generated using an explicit statistical error model:

$$o_{v,t} = m_{v,t} + S_v(\beta_v + \epsilon_{v,t}), \qquad \epsilon_{v,t} \sim N(0, \tau_v^2)$$

Where:
* $o_{v,t}$ is the simulated observation for variable $v$ at time $t$;
* $m_{v,t}$ is the ensemble mean prediction;
* $S_v$ is the scale factor (standard deviation of the ensemble mean);
* $\beta_v$ is the relative bias parameter (e.g. $0.30$, representing relative bias of $30\%$);
* $\tau_v$ is the relative random error standard deviation (e.g. $0.50$, representing random noise equal to $50\%$ of variability).

### R Simulation Code (Reference)

```r
# Side-note: R snippet used to generate observation fixtures from an error model
set.seed(42)
output_scale <- stats::sd(ensemble_mean)

relative_bias  <- 0.3 # 30% relative bias
relative_error <- 0.5 # 50% relative error noise

simulated_observation <- ensemble_mean +
  output_scale * (
    relative_bias +
      stats::rnorm(length(ensemble_mean), mean = 0, sd = relative_error)
  )
```

In `PEcAn.benchmark`, precomputed text fixtures (`small_salinas_ensemble.csv` and `small_salinas_obs.csv`) are provided in package `extdata` for fast, reproducible offline testing.

---

## 3. Data Intake, Reshaping, and Alignment

We load the package and read the precomputed fixtures:

```{r load_data}
library(PEcAn.benchmark)

# Path to built-in simulation fixtures
extdata_dir <- system.file("extdata", package = "PEcAn.benchmark")
if (extdata_dir == "" || !file.exists(file.path(extdata_dir, "small_salinas_ensemble.csv"))) {
  extdata_dir <- file.path(getwd(), "../inst/extdata")
}

model_csv <- file.path(extdata_dir, "small_salinas_ensemble.csv")
obs_csv   <- file.path(extdata_dir, "small_salinas_obs.csv")

raw_model <- read.csv(model_csv)
raw_obs   <- read.csv(obs_csv)

head(raw_model[, c("datetime", "site_id", "parameter", "variable", "prediction")])
head(raw_obs[, c("date", "site_id", "variable", "obs_mean", "obs_sd")])
```

### Reshaping EFI Long Data to Ensemble Matrix (`efi_long_to_array`)

To retain individual member trajectory information for spaghetti plots and probabilistic metrics (like CRPS), we reshape the long model dataset into an ensemble matrix:

```{r reshape}
# Reshape EFI long format dataframe to a (time x member) matrix
ens_mat <- efi_long_to_array(raw_model, var = "TotSoilCarb", site = "socs_sys1")

# Summary data frame with ensemble mean and 90% prediction intervals
model_summary <- data.frame(
  time = attr(ens_mat, "time"),
  value = rowMeans(ens_mat, na.rm = TRUE),
  model_q05 = apply(ens_mat, 1, quantile, probs = 0.05, na.rm = TRUE),
  model_q95 = apply(ens_mat, 1, quantile, probs = 0.95, na.rm = TRUE),
  site = "socs_sys1"
)

# Observation data frame
obs_filtered <- raw_obs[raw_obs$site_id == "socs_sys1" & raw_obs$variable == "TotSoilCarb", ]
obs_summary <- data.frame(
  time = as.POSIXct(obs_filtered$date, tz = "UTC"),
  value = obs_filtered$obs_mean,
  obvs_sd = obs_filtered$obs_sd,
  site = "socs_sys1"
)
```

### Temporal Alignment (`align_by_time`)

We align model predictions and observations in time within a half-year tolerance window (for annual observation alignment):

```{r align}
# Align model summary and observations by timestamp
aligned_df <- align_by_time(model_summary, obs_summary, tolerance_secs = 365 * 86400 / 2)

# Match ensemble matrix indices to aligned time points
aligned_indices <- match(aligned_df$time, model_summary$time)
aligned_ens_mat <- ens_mat[aligned_indices, , drop = FALSE]
attr(aligned_df, "ensemble_matrix") <- aligned_ens_mat

head(aligned_df)
```

---

## 4. Metric Calculations and Interpretation

The framework categorizes evaluation metrics into two groups:

1. **Point Metrics on Ensemble Mean**: Evaluates deterministic skill of the ensemble mean vector $\bar{m}$.
   * **BIAS / Mean Bias Error**: $\frac{1}{N}\sum (\bar{m}_i - o_i)$. Quantifies systematic over- or under-prediction.
   * **RMSE (Root Mean Squared Error)**: $\sqrt{\frac{1}{N}\sum (\bar{m}_i - o_i)^2}$. Overall magnitude of error.
   * **MAE (Mean Absolute Error)**: $\frac{1}{N}\sum |\bar{m}_i - o_i|$. L1 norm error distance.
   * **R2 (Squared Pearson Correlation)**: Coefficient of determination indicating proportion of variance explained.

2. **Spread / Probabilistic Metrics on Full Ensemble**: Evaluates prediction uncertainty and ensemble dispersion.
   * **COVERAGE (Prediction Interval Coverage)**: Fraction of observations falling within the 90% ensemble prediction interval $[q_{05}, q_{95}]$.
   * **CRPS (Continuous Ranked Probability Score)**: Evaluates full predictive probability distribution against observations using `scoringRules::crps_sample()` (or an analytical sample score fallback).

```{r metrics}
# Compute complete metric suite
metrics_results <- compute_metrics(
  aligned_df,
  metrics = c("BIAS", "RMSE", "MAE", "R2", "COVERAGE", "CRPS")
)

print(metrics_results)
```

---

## 5. Visualizations and Automated Reporting

### Spaghetti + Ribbon Timeseries Plot (`metric_timeseries_plot`)

Renders individual member trajectories (spaghetti) under the $q_{05}/q_{95}$ prediction ribbon alongside observations and error bars:

```{r plot_timeseries}
p_ts <- metric_timeseries_plot(aligned_df, var = "TotSoilCarb (gC/m2)")
print(p_ts)
```

### Diagnostic Scatter Plot (`metric_scatter_plot`)

Plots observed values against predicted ensemble mean values:

```{r plot_scatter}
p_sc <- metric_scatter_plot(aligned_df, var = "TotSoilCarb")
print(p_sc)
```

### Residual Plot (`metric_residual_plot`)

Plots residuals $(o_i - \bar{m}_i)$ against time or predicted values:

```{r plot_residual}
p_res <- metric_residual_plot(aligned_df, var = "TotSoilCarb")
print(p_res)
```

---

## 6. Substituting Custom Processed Data

To evaluate your own model predictions and observations:
1. Format model predictions in EFI long format CSV (`datetime`, `site_id`, `parameter`, `variable`, `prediction`).
2. Format observations with `date`/`time` and `obs_mean`/`obvs`.
3. Use `efi_long_to_array()` to extract the ensemble matrix.
4. Pass your aligned data into `align_by_time()` and `compute_metrics()`.
