Meta-analysis: Example of Single.MA
========================================================
Individual Meta-analysis

Individual meta-analysis for a specific trait and PFT is run by the function
single.MA. This will allow power analysis  to run repeated MA outside of the
full loop over traits and PFTs. 

The first step is to convert the R distribution to JAGS distribution. The prior transformed is then stored in jagsprior and the parameters are renamed as "distn", "a", "b" and "n".

```{r}
library(devtools)
load_all("~/pecan1/modules/meta.analysis")
load("~/pecan1/modules/meta.analysis/inst/extdata/singleMAvignetteinputs.RData")
jagsprior <- r2bugs.distributions(prior)
jagsprior <- jagsprior[, c('distn', 'parama', 'paramb', 'n')]
colnames(jagsprior) <- c("distn", "a", "b", "n")
colnames(prior)     <- c("distn", "a", "b", "n")
```
In order to determine how many levels of the factors to include in meta-analysis, the number of unique values of the greenhouse(ghs), site and trait (trt) variables are calculated. 
If one of the factors has more than 1 level, it will be included in the model.

```{r}
 # determine what factors to include in meta-analysis
    model.parms <- list(ghs  = length(unique(data$ghs)),
                        site = length(unique(data$site)),
                        trt  = length(unique(data$trt)))
    # define regression model
    reg.parms   <- list(ghs  = 'beta.ghs[ghs[k]]', #beta.o will be included by default
                        site = 'beta.site[site[k]]',
                        trt  = 'beta.trt[trt[k]]')
    if(sum(model.parms>1)==0) {
      reg.model <- ''
    } 
######################if(sum(model.parms>1)!=0)
    else {
      reg.model <- paste('+', reg.parms[model.parms > 1], collapse = " ")
    }

```
After deciding the levels of the factors, list of parameters will be generated for jags to follow.
The vairables "beta.o" and "sd.y" will always be considered. For greehouse (ghs) or trait (trt) factors, if there are p levels (p>1), (p-1) variables will be generated for it. For site (site),if there are p levels (p>1), p variables for each level will be generated. Only factors with more than one levels remain.
```{r}
vars <- c( 'beta.o', 'sd.y') 
    for (x in c('ghs', 'site', 'trt')) {
      if(model.parms[[x]] == 1) {
        data <- data[, which(names(data) != x)]
      } 
      else {
        data <- data
        if(x!='ghs') {
          vars <- c(vars, paste('sd.', x, sep = ''))
        }
        # m <- min(model.parms[[x]], 5)
        m <- model.parms[[x]]
        for (i in 1:m) {
          if(i == 1 && x == 'site') {
            vars <- c(vars, 'beta.site[1]')
          }
          if (i > 1) {
            vars <- c(vars, paste('beta.', x, '[', i, ']', sep=''))
          }
        }
      }
    }

```
The default JAGS model file is imported and modified based on user settings and default bug file.
```{r}
  modelfile = system.file("inst/ma.model.template.bug", package="PEcAn.MA")
    
    write.ma.model (modelfile = modelfile,
                    outfile = jag.model.file,
                    reg.model = reg.model,
                    jagsprior$distn, jagsprior$a, jagsprior$b,
                    n     = length ( data$Y ),
                    trt.n = model.parms[['trt']],
                    site.n= model.parms[['site']],
                    ghs.n = model.parms[['ghs']],
                    tauA  = tauA,
                    tauB  = tauB)

```

To overdisperse the n MCMC chains, intial values of beta.0 are set to the 1/(n+1),...,n/(n+1) quantiles of the prior on beta.o; for the default n=4 chains, this would be the {0.2, 0.4, 0.6, 0.8} quantiles
```{r}
 if(overdispersed == TRUE){
      ## overdispersed chains
      j.inits <- function(chain) list("beta.o" = do.call(paste('q',prior$dist,sep=''),
                                        list(chain * 1/(j.chains + 1), prior$a, prior$b)),
                                      .RNG.seed = chain,
                                      .RNG.name = "base::Mersenne-Twister")
    } else if (overdispersed == FALSE) {
      ## chains fixed at data mean - used if above code does not converge,
      ## invalidates assumptions about convergence, e.g. Gelman-Rubin diagnostic
      j.inits <- function(chain) list("beta.o" = mean(data$Y))
    }
    
```
Using the jags.model function, the Bayesian graphical model is generated based on the initials. The posterior samples in mcmc.list format are then created.
```{r}
 j.model   <- jags.model (file = jag.model.file,
                             data = data,
#                             n.adapt = 100, #will burn in below
                             inits = j.inits,
                             n.chains = j.chains)


    jags.out   <- coda.samples ( model = j.model,
                                variable.names = vars,
                                n.iter = j.iter,
                                thin = max(c(2,j.iter/(5000*2))))
return(jags.out)

```



