<<echo=false,results=hide>>=
  #options(warn=-1)

```{r}
library(devtools)
document("modules/meta.analysis")
test("modules/meta.analysis")
library(PEcAn.utils)
#outdir <- '~/archives/manuscript/out/ebifarm/'
#outdir <- 'F:/study/pecan/archives/manuscript/out/ebifarm/'

iter <- 10000
set.seed(0)
######################################
## Load results from prior runs
######################################
load(system.file("inst/extdata/prior","samples.Rdata",package="PEcAn.MA"))
prior.trait.samples <- trait.samples[['ebifarm.pavi']]

load(system.file("inst/extdata/prior","output.Rdata",package="PEcAn.MA"))
#load(paste(outdir, 'prior/output.Rdata', sep = ''))
prior.ensemble.output <- unlist(ensemble.output)[!is.na(unlist(ensemble.output))]

load(system.file("inst/extdata/prior","sensitivity.results.Rdata",package="PEcAn.MA"))
#load(paste(outdir, 'prior/sensitivity.results.Rdata', sep = ''))
prior.sensitivity.results <- sensitivity.results[[1]]

######################################
## Remove raw results from prior runs
######################################
rm(list = c("sensitivity.results", "ensemble.output", "trait.samples"))

####################################
# Load Results from Posterior runs 
####################################
load(system.file("inst/extdata/post","prior.distns.Rdata",package="PEcAn.MA"))
#load(paste(outdir, 'post/prior.distns.Rdata', sep = ''))
load(system.file("inst/extdata/post","trait.data.Rdata",package="PEcAn.MA"))
#load(paste(outdir, 'post/trait.data.Rdata', sep = ''))

## MCMC chains have already been truncated and thinned
load(system.file("inst/extdata/post","trait.mcmc.Rdata",package="PEcAn.MA"))
#load(paste(outdir, 'post/trait.mcmc.Rdata', sep = ''))

load(system.file("inst/extdata/post","samples.Rdata",package="PEcAn.MA"))
#load(paste(outdir, 'post/samples.Rdata', sep = ''))
post.ensemble.samples <- ensemble.samples[['ebifarm.pavi']]
post.trait.samples <- trait.samples[['ebifarm.pavi']]

### not sure why above does not have posteriors, but will create workaround
### needs to be fixed in R/write.configs.utils.R 
.tmp <- names(trait.mcmc)
for (i in .tmp) {
  post.trait.samples[[i]] <- as.matrix(trait.mcmc[[i]])[,'beta.o']
}

load(system.file("inst/extdata/post","output.Rdata",package="PEcAn.MA"))
#load(paste(outdir, 'post/output.Rdata', sep = ''))
post.ensemble.output  <- unlist(ensemble.output)[!is.na(unlist(ensemble.output))]

load(system.file("inst/extdata/post","sensitivity.results.Rdata",package="PEcAn.MA"))
#load(paste(outdir, 'post/sensitivity.results.Rdata', sep = ''))
post.sensitivity.results  <- sensitivity.results[[1]]


#########################################
## Remove raw results from posterior runs
#########################################
rm(list = c("sensitivity.results", "ensemble.output", "trait.samples", "ensemble.samples"))


## get Wang 2010 pavi yield data
paviyield <- read.csv(system.file("inst/extdata","wang2010data.csv",package="PEcAn.MA"), sep = ';', header = F)
colnames(paviyield) <- c('site', 'cultivar', 'ecotype', 'yield')

## All traits considered
traits <-  rownames(prior.distns) # c('mort2','growth_resp_factor','leaf_turnover_rate','leaf_width','nonlocal_dispersal','fineroot2leaf','root_turnover_rate','seedling_mortality','stomatal_slope','r_fract','root_respiration_rate','Vm_low_temp','SLA','Vcmax','quantum_efficiency')

## Traits included in meta-analysis
ma.traits <- names(trait.mcmc)  # c('SLA','leaf_width','seedling_mortality','Vcmax','fineroot2leaf','root_respiration_rate','root_turnover_rate')
```

```{r}

#subsection*{Trait Meta-analysis}
#<<trait.data,echo=FALSE,results=hide>>=
  sample.size <- lapply(trait.data, nrow)
```

Switchgrass data were collected from the literature and field for seven of the model parameters: specific leaf area (SLA) ($n=\Sexpr{sample.size$SLA}$), leaf width ($\Sexpr{sample.size$leaf_width}$), V$_{\text{cmax}}$ ($\Sexpr{sample.size$Vcmax}$), fine root to leaf allocation ratio ($\Sexpr{sample.size$fineroot2leaf}$), stomatal slope ($\Sexpr{sample.size$stomatal_slope}$), root respiration rate ($\Sexpr{sample.size$root_respiration_rate}$), and root turnover rate ($\Sexpr{sample.size$root_turnover_rate}$).
Table \ref{tab:maresults} summarizes the meta-analysis for each of these parameters, including the posterior mean and $95\%$ CI of the global mean, the fixed greenhouse effect, and each of the variance components (reported as standard deviations).
<<maresults,echo=FALSE,results=hide>>=
```{r}
  sla.ghs.stats <- tabnum(summary(trait.mcmc$SLA)$statistics["beta.ghs[2]",][c('Mean', 'SD')], 2)
######################
## Meta Analysis Table 
######################
mcmc.summaries <- lapply(trait.mcmc, summary)

## Make a table row

#To load the function get.stats.mcmc
#detach("package:PEcAn.utils", unload=TRUE)
#add @export in the util.R file infront of the get.stats.mcmc function and save
#document("utils")

ma.results.list <- lapply(ma.traits, 
                          function(x) get.stats.mcmc(mcmc.summaries[[x]], 
                                                     sample.size[[x]]))
ma.results <- t(as.data.frame(ma.results.list))

rownames(ma.results) <- NULL
#trait.dictionary <- read.csv(system.file("data","trait.dictionary.csv",package="PEcAn.utils"), sep = ';', header = F)
#make some modification for the following code: 
#ma.results <- data.frame(Variable = trait.dictionary(ma.traits)$figid, ma.results)
ma.results <- data.frame(Variable = ma.traits, ma.results)
colnames(ma.results) <- c("Variable", 'n', '$\\beta{_0}$',"$\\sigma_{Y}$", '$\\sigma_{\\text{site}}$', '$\\sigma_{\\text{treatment|site}}$', '$\\beta_{\\text{greenhouse}}$')

```
<<results=hide,echo=FALSE,label=matable>>=
```{r}
  print(xtable(ma.results, 
               caption = "{\\bf Meta-analysis Results} Results of meta-analysis of Switchgrass data for six physiological traits.  The global mean parameter, $\\beta{_0}$, is used to parametrize the Ecosystem Demography model and is described in more detail by Figure~\\ref{fig:traitpdfs}. The variance components are transformed from precision to the standard deviation scale for ease of interpretation. Values are reported as the parameter median with the $95\\%$ CI in parentheses. Units are the same as in Table~\\ref{tab:priors}.",
               caption.placement = 'top',
               align = "lrllllll",
               label = "tab:maresults"), 
        file = 'maresultstab.tex',
        floating.environment = "sidewaystable",
        #      table.placement = "ht",
        sanitize.colnames.function = identity,
        sanitize.text.function = identity,
        include.rownames = FALSE,)
```
@ 
<<results=hide,echo=FALSE,label=mafig>>=
  #######################
## Meta Analysis Figure 
#######################
plot.trait.data <- trait.data[ma.traits]
plot.trait.samples <- post.trait.samples[ma.traits]

trait.plot.fn <- function(x, ...){
  plot.trait(trait = x, 
             prior = prior.distns[x, 1:3], 
             posterior.sample = post.trait.samples[[x]],
             trait.df = plot.trait.data[[x]], ...)
}

plot.args <- list('seedling_mortality' = 
                    list(x.lim = c(0.1, 1), 
                         y.lim = c(0,50)
                    ),#,logx = TRUE),
                  'fineroot2leaf' = list(x.lim = c(0,10)),
                  'SLA' = list(x.lim = c(0,40)),
                  'root_turnover_rate' = list(x.lim = c(0,2)))


trait.plots <- list()

pdf('traitpdfsi.pdf', width = 6, height = 3)
for(x in rownames(prior.distns)[rownames(prior.distns) %in% ma.traits]){
  if(x %in% ma.traits){
    p <- do.call(trait.plot.fn, 
                 append(list(x = x, 
                             fontsize = list(title = 24, axis = 18)), 
                        plot.args[[x]]))
    trait.plots[[x]] <- p
    print(p)
  }
}
dev.off()
system("pdflatex traitpdfs.tex", ignore.stdout = TRUE)

@ 

<<results=hide,echo=false>>=
  sla.ghs <- as.matrix(trait.mcmc$SLA)[,"beta.ghs[2]"]
p.slaghs <- 2*sum(sla.ghs < 0) / length(sla.ghs)
lw.ghs <- as.matrix(trait.mcmc$leaf_width)[,"beta.ghs[2]"]
p.lwghs <- 2*sum(lw.ghs < 0)/length(lw.ghs)

sla.s <- paste("$",gsub("\\$","",
                        paste(colnames(ma.results)[4],'_\\text{SLA}=', ma.results[1,4], ',' ,
                              colnames(ma.results)[5],'_\\text{SLA}=',ma.results[1,5], 
                              sep = "")),"$")
lw.s  <- paste("$",gsub("\\$","",
                        paste(colnames(ma.results)[4],'_\\text{lw}=', ma.results[2,4], ',' ,
                              colnames(ma.results)[5],'_\\text{lw}=',ma.results[2,5], 
                              sep = "")),"$")

@ 

SLA and leaf width data were from from multiple sites, but the meta-analysis provided no evidence for among site variability in excess of within site variability ($\sigma_Y$ and  $\sigma_{\text{site}}$, respectively, in Table \ref{tab:maresults}).  
For the remaining traits, there was insufficient spatial sampling to assess site to site variability.
Greenhouse growing conditions had a positive effect on both SLA ($P  = \Sexpr{tabnum(p.slaghs,2)}$), and leaf width ($P = \Sexpr{tabnum(p.lwghs,2)})$.

Figure \ref{fig:traitpdfs} compares parameters before and after incorporating data in the meta-analysis.
A reduction in parameter uncertainty is seen as the reduction in the spread of the posterior (black) compared to the prior (grey) parameter distributions.
The influence of the prior information on the posterior distribution increased when the prior was more constrained or when less data were available for use in the meta-analysis.
For example, data substantially constrained the uncertainty in the V$_{\text{cmax}}$ and SLA posteriors relative to the priors.
By contrast, there was little effect of additional data on the parameter estimates for fine root to leaf allocation and root respiration rate; these parameters had relatively well constrained priors and limited species-specific data.
% Burning treatment doubled leaf width, N fert had no effect on Vcmax. 
% treatment effects on SLA were all in ghs, mostly w/ fertilization
% select treatments.name, treatments.control, sites.greenhouse, sites.id, traits.mean from treatments join traits on traits.treatment_id = treatments.id join sites on traits.site_id = sites.id where traits.specie_id = 938 and variable_id in (select id from variables where name = "Vcmax") order by sites.id, treatments.id;


\subsection*{Model Analysis}

<<label=model.analysis,results=hide,echo=FALSE>>=
  ## ## To look at the dead vs. live plants
  ## hist(c(prior.ensemble.output, post.ensemble.output), breaks = 0:440/4, xlim = c(0,3))
  ## hist(c(prior.ensemble.output, post.ensemble.output), breaks = 0:440/4)
  
  #### Comparison of percentage dead in prior vs posterior runs
  #### Predict that more die in prior runs 
  ####    because there are more unrealistic parameter combinations
  prior.is.dead <- prior.ensemble.output < 2
post.is.dead  <-  post.ensemble.output < 2

## summarize p(dead) for posterior and prior
prior.dead <- prior.ensemble.output[prior.is.dead]
prior.live <- prior.ensemble.output[!prior.is.dead]
post.dead  <- post.ensemble.output[post.is.dead]
post.live  <- post.ensemble.output[!post.is.dead]

#### Test to see if means are different
## mu prior
mu <- 0
## sigma prior
s  <- 100

### Following Albert 2007 p56-57, evaluate posterior distribution of live and dead plants
##    assume standard noninformative prior: g(mu, 1/sigma2) \propto 1/sigma2
normal.posterior <- function(x, iter = 5000){
  SS <- sum((x - mean(x))^2)
  n  <- length(x)
  sigma2 <- SS/rchisq(iter, n - 1)
  mu <- rnorm(iter, mean = mean(x), sd = sqrt(sigma2)/sqrt(n))
  return(list(mu = mu, sigma2 = sigma2))
}

prior.dead.posterior <- normal.posterior(prior.dead)
post.dead.posterior <- normal.posterior(post.dead)
prior.live.posterior <- normal.posterior(prior.live)
post.live.posterior <- normal.posterior(post.live)

#### test differences between post / prior for dead and live 
dead.diff <- prior.dead.posterior$mu - post.dead.posterior$mu
live.diff <- prior.live.posterior$mu - post.live.posterior$mu

quantile(dead.diff, c(0.025, 0.5, 0.975))
quantile(live.diff, c(0.025, 0.5, 0.975))

#### Test to see if proportion of dead are different 
## using ad-hoc simulation from beta-binomial posterior
## beta priors (~beta(1,1))
a1 <- a2 <- b1 <- b2 <- 1
## binomial likelihood parameters
prior.dead.n <- sum(prior.is.dead, na.rm = TRUE)
prior.n = sum(!is.na(prior.is.dead))
post.dead.n <- sum(post.is.dead, na.rm = TRUE)
post.n = sum(!is.na(post.is.dead))

## posteriors of beta-binomial
prior.pdead.posterior <- rbeta(10000, prior.dead.n + a1, prior.n - prior.dead.n + b1)
post.pdead.posterior  <- rbeta(10000, post.dead.n  + a2, post.n  - post.dead.n  + b2)

## summary
prior.pdead.summary <- quantile(prior.pdead.posterior, c(0.025, 0.5, 0.975)) 
post.pdead.summary  <- quantile(post.pdead.posterior, c(0.025, 0.5, 0.975))
## means comparison
dead.diff     <- prior.pdead.posterior - post.pdead.posterior
dead.diff.test <- quantile(dead.diff, c(.025, 0.5, .975))
dead.alpha <- ecdf(dead.diff)(0)/2 ## significance level

prior.stats <- tabnum(quantile(prior.live, c(0.1, 0.5, 0.9)))
post.stats  <- tabnum(quantile(post.live, c(0.1, 0.5, 0.9)))
prior.var   <- tabnum(var(prior.live))
post.var    <- tabnum(var(post.live))

#### calculate variance estimates for posterior runs

############################
### Ensemble Runs (n = 500)
############################
splinefuns <- post.sensitivity.results$sensitivity.output$sa.splines

## Summary stats
mean.f.Phi   <- mean(post.ensemble.output)
median.f.Phi <- median(post.ensemble.output)
## var(f(Phi)): total ensemble variance 
var.f.Phi   <- variance.stats(post.ensemble.output)

## spline estimate of ensemble
## Sensitivity Analysis only conducted one run at median values
##    so the median row of output is all from the same run, will take the first
gi.phii.median    <- sensitivity.output[['ebifarm.pavi']]["50",1]

## Calculate spline ensemble for 500 parameter sets used in model
require(plyr)
gi.phii.500 <- get.gi.phii(splinefuns, 
                           post.ensemble.samples) 
spline.ensemble.500 <- spline.ensemble(gi.phii.500, 
                                       median = gi.phii.median)

spline.var.500      <- variance.stats(spline.ensemble.500)
############################
### Ensemble Runs (n = 10000)
############################

gi.phii.10k <- get.gi.phii(splinefuns, 
                           post.trait.samples, 
                           maxn = 10000)

spline.ensemble.10k <- spline.ensemble(gi.phii.10k, median = gi.phii.median)
spline.var.10k      <- variance.stats(spline.ensemble.10k)
spline.dead.logical <- spline.ensemble.10k < 2 
spline.live <- spline.ensemble.10k[!spline.dead.logical]
spline.dead <- spline.ensemble.10k[spline.dead.logical]
spline.dead.n <- sum(spline.dead.logical)
spline.n <- length(spline.ensemble.10k)

@ 

\subsubsection*{Ensemble}
Within the model ensemble analysis (Figure \ref{fig:ensembledensity}), both the prior and posterior parameterizations produced yield estimates that were consistent with yields observed at the Urbana site for which the model was run \citep{heaton2008mub} and with $1902$ previously reported yields of switchgrass \citep{wang2010qrc}.
In both the prior and posterior ensembles, the predicted aboveground biomass was clearly bimodal.
These two modes had little overlap and a distinct break at two Mg/ha.
We infered that the first peak represents non-viable plants generated by unrealistic parameter sets so plants with aboveground biomass less than two Mg/ha were considered ``non-viable''.
When summarizing the model output, we consider viable and non-viable ensemble members separately;  all runs are considered in the senstivity analysis and variance decomposition.
A greater percentage of runs in the prior ensemble fell below this threshold ($\Sexpr{tabnum(prior.pdead.summary[2]*100)}$ vs $\Sexpr{tabnum(post.pdead.summary[2]*100)}$, $P \simeq \Sexpr{signif(dead.alpha, 2)}$).

Compared to the prior ensemble prediction, the data-constrained posterior runs had lower median yields and a more constrained $95\%$ credible interval ($\Sexpr{post.stats[2]} [\Sexpr{tabnum(post.stats[1],2)}, \Sexpr{tabnum(post.stats[3],2)}]$ Mg/ha vs $\Sexpr{prior.stats[2]} [\Sexpr{tabnum(prior.stats[1],2)}, \Sexpr{tabnum(prior.stats[3],2)}]$ Mg/ha).
This reflects the substantial shrinkage of the posterior relative to the prior SD estimates of model output uncertainty (from $\sigma = \Sexpr{tabnum(sqrt(prior.var))}$ to $\sigma = \Sexpr{tabnum(sqrt(post.var))}$).
In particular, the upper tail of the modeled yield was reduced toward the observed yields.
Despite the reduction in ensemble uncertainty, the ensemble posterior yield was still relatively imprecise and had much greater uncertainty than the field trial \citep[][$\sigma = 4.1$]{heaton2008mub}  or the meta-analysis of all observations \citep[][$(\sigma = 5.4)$]{wang2010qrc}.
The spline ensemble viable plants had a median $\Sexpr{tabnum(median(spline.live))}[\Sexpr{tabnum(quantile(spline.live, 0.025),2)},\Sexpr{tabnum(quantile(spline.live, 0.975),2)}]$ and $\sigma=\Sexpr{tabnum(sd(spline.live), 2)}$.

\subsubsection*{Sensitivity Analysis}
Sensitivity analysis demonstrated that traits varied in their effect on on aboveground biomass (Figure \ref{fig:agbsa}), and many of these relationships are clearly non-polynomial.
For example, parameters associated with photosynthesis and carbon allocation - including V$_{\text{cmax}}$, SLA, growth respiration, and root allocation - were particularly sensitive.
For particularly sensitive parameters, the sensitivity functions had coverage of unrealistic yields greater than 30 Mg/ha.
Constraining SLA and V$_{\text{cmax}}$ parameters with data resulted in a more realistic range of yields.   
On the other hand, aboveground biomass was largely insensitive to leaf width, seed dispersal, and mortality rate.

<<results=hide,echo=FALSE,label=sensitivity.analysis>>=
  sa.plots <- plot.sensitivities(post.sensitivity.results$sensitivity.output,
                                 prior.sensitivity.results$sensitivity.output,
                                 linesize = 1,
                                 dotsize = 3,
                                 fontsize = list(title = 24, axis = 18))
pdf('sensitivityanalysis.pdf', width = 8, height = 4)
for(x in rownames(prior.distns)){
  print(sa.plots[[x]] + opts(panel.border = theme_border(c("bottom","left"))))
}
dev.off()
system('pdflatex sensitivityanalysisplots.tex', ignore.stdout = TRUE)
system('pdflatex sensitivityanalysisplotslabel.tex', ignore.stdout = TRUE)

@ 
\subsubsection*{Variance Decomposition}

<<results=hide,echo=false,label=variance.decomposition>>=
  
  #############################
## Variance Decomposition 
## (sum of variances associated with each trait)
#############################
vd.var.500 <- vd.variance(gi.phii.500)
vd.var.10k <- vd.variance(gi.phii.10k)

##########################################################
## variance decomposition plot (a.k.a. popsicle tryptich)
##########################################################

vd.plots <-  plot.variance.decomposition(post.sensitivity.results$variance.decomposition.output,
                                         prior.sensitivity.results$variance.decomposition.output)

pdf('variancedecomposition.pdf', width = 14, height = 8)
do.call(grid.arrange, c(vd.plots, ncol = 4))
dev.off() 

@

<<echo=FALSE,results=hide,label=ensemble.plots>>=
  
  font.adjust <-  1.5
## plot parameters
theme_set(theme_bw())
opt <- opts(title = "Modeled and Observed Switchgrass Yield",
            axis.text.y = theme_blank(),#theme_text(size = 14 * font.adjust),
            axis.text.x = theme_text(size = 14 * font.adjust),
            axis.title.x = theme_text(size = 14 * font.adjust), 
            axis.title.y = theme_text(size = 14 * font.adjust, angle = 90),
            plot.title = theme_text(size = 16 * font.adjust), 
            panel.grid.major = theme_blank(),
            panel.grid.minor = theme_blank(),
            panel.border = theme_border(c("bottom","left")),
            legend.position = "none",
            axis.color = 'black',
            plot.background = theme_blank()
)

xlim <- c(0,100)
live.breaks <- seq(0, 100, by = 2)
dead.breaks <- seq(0, 2,   by = 0.2)
x.ticks <- pretty(xlim)
y.ticks <- c(0:7)*20
x.breaks  <- c(seq(0,1.75, by = 0.25), seq(2,100, by = 2.5))

heaton.data <- data.frame(mean = c(11.5, 10.6, 15.6), sd = c(1.8, 1.3, 2.6) * sqrt(4))
### there is a closed form solution to this called the 'normal-normal' model
### Lynch, S. M. (2007). Introduction to Applied Bayesian Statistics and Estimation for Social Scientists. New York: Springer.
heaton.sim <- c(rnorm(10000, 11.5, 1.8*sqrt(3)), rnorm(10000, 10.6, 1.3 * sqrt(3)), rnorm(10000, 15.6, 2.6 * sqrt(3))) 
heaton.median <- median(heaton.sim)
heaton.ci <- quantile(heaton.sim, pnorm(c(-1.96, -1, 1, 1.96)))

wang.median <- median(paviyield$yield)
wang.ci   <- quantile(paviyield$yield, pnorm(c(-1.96, -1, 1, 1.96)))

post.live.median <- median(post.live)
post.live.ci <- quantile(post.live, pnorm(c(-1.96, -1, 1, 1.96)))
post.all.median <- median(post.ensemble.output)
post.all.ci <- quantile(post.ensemble.output, pnorm(c(-1.96, -1, 1, 1.96)))


prior.live.median <- median(prior.live)
prior.live.ci <- quantile(prior.live, pnorm(c(-1.96, -1, 1, 1.96)))
prior.all.median <- median(prior.ensemble.output)
prior.all.ci <- quantile(prior.ensemble.output, pnorm(c(-1.96, -1, 1, 1.96)))


## set height of Wang and Heaton data relative to y-axis
wang.y <- 100
heaton.y <- wang.y + 10## for density: wang.y - 0.01
prior.live.y   <- wang.y - 20
prior.all.y   <- wang.y - 22.5
post.live.y   <- wang.y - 32.5
post.all.y   <- wang.y - 35

base.plot <- ggplot() + opt 

ens.plot <-  base.plot +   
  ## shaded area for non-viable plants
  geom_rect(aes(xmin = 0, xmax = 2, ymin = 0, ymax = 150), 
            fill = "grey95", linetype = 'dashed', color = 'grey60') + 
  geom_text(aes(x = 1.5, y = 110, label = 'non-viable plants'), 
            angle = 90, color = "grey60", size = 4 * font.adjust) +
  ## Spline ensemble
  geom_step(aes(x = spline.ensemble.500), 
            stat = 'bin', size = 0.6, breaks = x.breaks - 0.1, 
            color = 'grey50', linetype = 'dotted') + 
  geom_text(aes(x = 35, y = 36), label = 'spline ensemble', 
            hjust = 0, color = "grey50", size = 4 * font.adjust) +
  geom_segment(aes(x = 35, y = 36, xend = 30.8, yend = 22), 
               color = "grey45", linetype = 'dotted') +
  ## ED2 Prior ensemble hist
  geom_step(aes(x = prior.ensemble.output), 
            stat = 'bin', size = 0.6, breaks = x.breaks + 0.1, 
            limits = c(0,100), color = 'grey45') +   
  geom_text(aes(x = 30, y = 44), label = 'ED2 prior ensemble', 
            hjust = 0, color = "grey45", size = 4 * font.adjust) +
  geom_segment(aes(x = 30, y = 44, xend = 20, yend = 17), 
               color = "grey45") +
  ## ED2 Ensemble hist
  geom_step(aes(x = post.ensemble.output), 
            stat = 'bin', size = 0.6, breaks = x.breaks) +   
  geom_text(aes(x = 25, y = 52), label = 'ED2 ensemble', hjust = 0, size = 4 * font.adjust) +
  geom_segment(aes(x = 25, y = 52, xend = 19, yend = 36)) +
  ## Wang 2008 hist.
  geom_step(data = paviyield, 
            aes(x=yield, y = ..density.. * 500), 
            stat = 'bin', breaks = seq(0, 40, by = 2.5), 
            linetype = 'dashed') +  
  geom_text(aes(x = 40, y = 28), label = 'Wang et al. 2010', hjust = 0, size = 4 * font.adjust) +
  geom_segment(aes(x = 40, y = 28, xend = 32, yend = 0), 
               linetype = 'dashed') + 
  scale_x_continuous( "Aboveground Biomass, Mg/ha", expand = c(0, 0)) + 
  scale_y_continuous("density", limits = c(0,150), expand = c(0,0)) +
  ##Data from Wang et al 2010 as bar
  geom_segment(aes(x = wang.ci[1], 
                   xend = wang.ci[4], y = wang.y, yend = wang.y), 
               size = 0.4) + # 95%CI range
  geom_segment(aes(x= wang.ci[2], y = wang.y, xend = wang.ci[3],  yend = wang.y), size = 1) +
  geom_point(aes(x = wang.median, y = wang.y), size = 4, fill = 'black', shape = 5) + 
  geom_text(aes(x = 20, y = wang.y + 3), label = 'Wang et al. 2010, n = 1902', hjust = 0, size = 4 * font.adjust) +
  ## Data from Heaton et al 2008
  geom_segment(aes(x= heaton.ci[2], y = heaton.y, xend = heaton.ci[3],  yend = heaton.y), size = 1) + # mean +/- SD
  geom_segment(aes(x= heaton.ci[1], y = heaton.y, xend = heaton.ci[4],  yend = heaton.y), size = 0.4) + # 95%CI
  geom_point(aes(x = heaton.median, y = heaton.y), size = 4, fill = 'black', shape = 5) + 
  geom_text(aes(x = 20, y = heaton.y + 3), label = 'Heaton et al. 2008, n = 4', hjust = 0, size = 4 * font.adjust) +
  ## Model posterior live
  geom_segment(aes(x = post.live.ci[2], y = post.live.y, xend = post.live.ci[3],  yend = post.live.y), size = 1) + # mean +/- SD
  geom_segment(aes(x= post.live.ci[1], y = post.live.y, xend =post.live.ci[4],  yend = post.live.y), size = 0.4) + # 95%CI
  geom_point(aes(x = post.live.median, y = post.live.y), size = 4, fill = 'black', shape = 5) + 
  geom_text(aes(x = 20, y = post.live.y + 5), label = 'ED2 ensemble', hjust = 0, size = 4 * font.adjust) + 
  ## Model posterior all
  #  geom_segment(aes(x = post.all.ci[2], y = post.all.y, xend = post.all.ci[3],  yend = post.all.y), size = 1) + # mean +/- SD
  #  geom_segment(aes(x= post.all.ci[1], y = post.all.y, xend =post.all.ci[4],  yend = post.all.y), size = 0.4) + # 95%CI
  #  geom_point(aes(x = post.all.median, y = post.all.y), size = 4, fill = 'black', shape = 5) + 
  #  geom_text(aes(x = 50, y = post.live.y + 0.6 ), label = 'viable plants', hjust = 0, vjust = 0, size = 4 * font.adjust) + 
  #  geom_text(aes(x = 35, y = post.all.y - 2.1), label = 'all', hjust = 0, vjust = 0, size = 4 * font.adjust) +
  ## Model prior live
  geom_segment(aes(x = prior.live.ci[2], y = prior.live.y, xend = prior.live.ci[3],  yend = prior.live.y), size = 1, color = "grey45") + # mean +/- SD
  geom_segment(aes(x= prior.live.ci[1], y = prior.live.y, xend =prior.live.ci[4],  yend = prior.live.y), size = 0.4, color = "grey45") + # 95%CI
  geom_point(aes(x = prior.live.median, y = prior.live.y), size = 4, fill = 'black', shape = 5, color = "grey45") + 
  geom_text(aes(x = 20, y = prior.live.y + 5), label = 'ED2 prior ensemble', hjust = 0, color = "grey45", size = 4 * font.adjust) + 
  ## Model prior all
  #  geom_segment(aes(x = prior.all.ci[2], y = prior.all.y, xend = prior.all.ci[3],  yend = prior.all.y), size = 1, color = "grey45") + # mean +/- SD
  #  geom_segment(aes(x= prior.all.ci[1], y = prior.all.y, xend =prior.all.ci[4],  yend = prior.all.y), size = 0.4, color = "grey45") + # 95%CI
  #  geom_point(aes(x = prior.all.median, y = prior.all.y), size = 4, fill = 'black', shape = 5, color = "grey45") +
  #  geom_text(aes(x = 80, y = prior.live.y ), label = 'viable plants', hjust = 0, vjust = 0, size = 4 * font.adjust, color = "grey45") + 
  #  geom_text(aes(x = 35, y = prior.all.y - 2.5), label = 'all', hjust = 0, vjust = 0, size = 4 * font.adjust, color = "grey45") +
  ##  Fix Axis
  geom_segment(aes(x = 0, y = 0, xend = 100, yend = 0), color = 'black') +
  geom_segment(aes(x = 0, y = 0, xend = 0, yend = 150), color = 'black') 
pdf('ensembledensity.pdf', height = 8, width = 16)
ens.plot
dev.off()

@ 
<<label=vartable,results=hide,echo=false>>=
  
  print.varsummary <- function(varest){
    v <- tabnum(sqrt(varest$var),3)
    s <- tabnum(sqrt(varest$sd), 2)
    paste.sd.sd_sd <- paste(v, "(", s, ")", sep = '')
    return(paste.sd.sd_sd)
  }

variancetab <- data.frame("model" = c("ensemble",
                                      "$s_{f(\\vec{\\beta_0})}$",
                                      print.varsummary(var.f.Phi),
                                      "$^*$"),
                          "spline" = c("ensemble", 
                                       "$s_{\\vec{g}(\\vec{\\beta{_0}})}$", 
                                       print.varsummary(spline.var.500), 
                                       print.varsummary(spline.var.10k)),
                          "variance" = c("decomposition", 
                                         "$\\sum{s_{g_i(\\beta{_0}_i)}}$",
                                         print.varsummary(vd.var.500),
                                         print.varsummary(vd.var.10k)))
rownames(variancetab) <- c("", "$n$", "500", "10000")


print(xtable(x = variancetab, 
             label = "tab:variance",  
             align = "rccc"), 
      file = "variancetab.tex",
      floating = FALSE,
      #      floating.environment= "table",
      #      table.placement = "ht",
      hline.after = c(-1, 2, 4),
      include.rownames = TRUE, 
      sanitize.text.function = function(x) gsub("%", "\\\\%", x),
      sanitize.colnames.function = identity)

@ 

<<label=omegatable,results=hide,echo=false>>=
  
  ssgi <- sqrt(vd.var.10k$var)
sg   <- sqrt(spline.var.10k$var)
sf   <- sqrt(var.f.Phi$var)
omega <- (ssgi-sf)

omegatab <- data.frame(rbind(c("$\\sum{s_{g_i}} - s_f$", 
                               tabnum(ssgi - sf,2)),
                             c("$s_{\\vec{g}} - s_f$", 
                               format(sg - sf, digits = 1, nsmall = 1 )),
                             c("$\\sum{s_{g_i}} - s_{\\vec{g}}$", 
                               tabnum(ssgi - sg, 2))))

rownames(omegatab) <- c("$\\omega_{\\text{total}}$",
                        "$\\omega_{\\text{covariance}}$",
                        "$\\omega_{\\text{truncation}}$")
colnames(omegatab) <- c("calculation", "Mg/ha")
print(xtable(x =  omegatab, 
             label = "tab:omega",  
             align = "lrc"), 
      file = "omegatab.tex",
      floating = FALSE,
      #      floating.environment= "table",
      #      table.placement = "ht",
      hline.after = c(-1, 0, 3),
      include.rownames = TRUE, 
      sanitize.text.function = function(x) gsub("%", "\\\\%", x),
      sanitize.colnames.function = identity)

@ 

The variance decomposition showed that data-constrained parameters substantially reduced their contribution to overall model variance (Figure \ref{fig:vd}).
Prior to including species-specific trait data, SLA contributed the most to model uncertainty, followed by growth respiration, fine root allocation, V$_{\text{cmax}}$, seedling mortality, and stomatal slope (right panel, grey bars Figure \ref{fig:vd}).
Incorporating species level data substantially reduced the contributions of SLA, V$_{\text{cmax}}$, seedling mortality, and stomatal slope to model uncertainty.
For example, SLA fell from first to fourth and stomatal slope fell from sixth to fourteenth in rank contribution to ensemble variance.
Although the addition of data reduced parameter uncertainty for fine root to leaf allocation, aboveground biomass was more sensitive to this parameter at the posterior median. 
These changes cancelled each other out, and as a result the contribution of the fine root allocation parameter to ensemble variance remained constant.

The variance of the ensemble was less than the variance calculated in the variance decomposition, and this difference is the closure term, $\omega$. 
The closure term accounted for approximately $\Sexpr{omega/ssgi*100}\%$ of the variance decomposition estimate (Table~\ref{tab:omega}).
There was no effect of increasing the sample size from 500 to 10000 on the variance estimates (Table~\ref{tab:variance}).
