---
title: "Pre Release database Cleanup"
output: html_notebook
---

This is an quick script for cleaning up the database. For further documentation see the README.rmd file in the main `qaqc` folder. 

**Step 1: set up an outdir and a connection to bety**
The outdir is where temporary editing files will be written, and where a backup of bety will be stored. 
```{r,eval=FALSE}
con <- RPostgreSQL::dbConnect(RPostgreSQL::PostgreSQL(),
dbname = "bety",
password = 'bety',
host = 'psql-pecan.bu.edu',
user = 'bety')
outdir<-as.character(getwd()) #set as prefered out directory
bety_backup_directory<-as.character(getwd()) #set as prefered directory for bety backup.

options(scipen=999) #To view full id values
```

**Step 2: Back up bety**
Before any major deletion processes, it makes sense to make a back-up version of the database. Don't skip this step. 
```{r,eval=FALSE}
system('TODAY=$( date +"%d" )')
backup_dir<-paste('pg_dump -U bety -d bety | gzip -9 > ', bety_backup_directory,'/bety-pre-culling${TODAY}.sql.gz', sep="")

system(backup_dir)
```

**Step 3: find all of the entries that should be deleted ***
```{r,eval=FALSE}

formats<-find_formats_without_inputs(con=con, created_before = "2016-01-01", user_id = NULL, updated_after = "2016-01-01") #Ideally, date should be set to the date of the last release 

inputs<-find_inputs_without_formats(con=con, created_before = "2014-01-01",updated_after = NULL, user_id = NULL)

```

Since just a dump of every column can be hard to read, just choose the columns that are important. 
```{r,eval=FALSE}
column_names<-get_table_column_names(table = formats, con = con)
column_names$formats

column_names<-column_names$formats[column_names$formats %in% c("id","created_at", "name", "notes", "table_name")]
column_names
```

** Option 1: Edit an R object ***

This is the most important step! Navigate to the written out table and *delete entries that should remain in the database*. 
```{r,eval=FALSE}
formats<-formats[colnames(formats) %in% column_names] #subset for easy viewing
View(formats)

formats<-formats[grep("test",formats$name),]
View(formats)
```

** Option 2: Edit a file  **

This is also the most important step! Navigate to the written out table and *delete entries that should remain in the database*. If the tables are difficult to read, change what columns are retained by editing the "relevant_table_columns" parameter.

```{r,eval=FALSE}
write_out_table(table = formats, outdir = outdir, relevant_table_columns = column_names, table_name = "formats")
write_out_table(table = inputs,outdir = outdir, relevant_table_columns =c("id", "created_at", "name"), table_name = "inputs")
```

**Step 5: Cull all remaining entries in the files**

This will delete all entries remaining in the file, and place a delete log in the outdir directory. 

*please do not run before editing either the table object or the "query_of" file*
```{r}
### Note: Do not run before the editing step ###

## From an R object ##
#cull_database_entries(table=formats, outdir = outdir, con=con, table_name = "formats")

## From a file ##
#cull_database_entries(outdir = outdir,file_name ='query_of_formats', con=con, table_name = "formats")
#cull_database_entries(outdir = outdir,file_name ='query_of_inputs', con=con, table_name = "inputs")

```

