# The PEcAn Docker API

If you have a running instance of Dockerized PEcAn (or other setup where PEcAn workflows are submitted via [RabbitMQ](#rabbitmq)),
you have the option of running and managing PEcAn workflows using the `pecanapi` package.

For more details, see the `pecanapi` [package vignette](#pecanapi-vignette) and function-level documentation.
What follows is a lightning introduction.

## Installation

The package can be installed directly from GitHub via `devtools::install_github`:

```{r, eval = FALSE}
devtools::install_github("pecanproject/pecan/api@develop")
```

## Creating and submitting a workflow

With `pecanapi`, creating a workflow, submitting it to RabbitMQ, monitoring its progress, and processing its output can all be accomplished via an R script.

Start by loading the package (and the `magrittr` package, for the `%>%` pipe operator).

```{r, eval = FALSE}
library(pecanapi)
library(magrittr)
```

Set your PEcAn database user ID, and create a database connection object, which will be used for database operations throughout the workflow.

```{r, eval = FALSE}
options(pecanapi.user_id = 99000000002)
con <- DBI::dbConnect(
  RPostgres::Postgres(),
  user = "bety",
  password = "bety",
  host = "localhost",
  port = 5432
)
```

Find model and site IDs for the site and model you want to run.

```{r, eval = FALSE}
model_id <- get_model_id(con, "SIPNET", "136")
all_umbs <- search_sites(con, "umbs%disturbance")
site_id <- subset(all_umbs, !is.na(mat))[["id"]]
```

Insert a new workflow into the PEcAn database, and extract its ID.

```{r, eval = FALSE}
workflow <- insert_new_workflow(con, site_id, model_id,
                                start_date = "2004-01-01",
                                end_date = "2004-12-31")
workflow_id <- workflow[["id"]]
```

Pull all of this information together into a settings list object.

```{r, eval = FALSE}
settings <- list() %>%
  add_workflow(workflow) %>%
  add_database() %>%
  add_pft("temperate.deciduous") %>%
  add_rabbitmq(con = con) %>%
  modifyList(list(
    meta.analysis = list(iter = 3000, random.effects = FALSE),
    run = list(inputs = list(met = list(source = "CRUNCEP", output = "SIPNET", method = "ncss"))),
    ensemble = list(size = 1, variable = "NPP")
  ))
```

Submit the workflow via RabbitMQ, and monitor its progress in the R process.

```{r, eval = FALSE}
submit_workflow(settings)
watch_workflow(workflow_id)
```

Use THREDDS to access and analyze the output.

```{r, eval = FALSE}
sipnet_out <- ncdf4::nc_open(run_dap(workflow_id, "2004.nc"))
gpp <- ncdf4::ncvar_get(sipnet_out, "GPP")
time <- ncdf4::ncvar_get(sipnet_out, "time")
ncdf4::nc_close(sipnet_out)
plot(time, gpp, type = "l")
```
