---
title: "The PEcAn RTM package"
author: Alexey Shiklomanov
output_format: pdf_document
---

# Introduction
The PEcAnRTM package provides tools for analyses involving common radiative 
transfer models. The highlights of this package are its ability to efficiently 
run a suite of related leaf and canopy radiative transfer models, as well as to 
perform maximum likelihood and, particularly, Bayesian inversions of these 
models. 

# Installation
The easiest way to install this package is via `install_github` from the 
`devtools` package.

```{r, eval=FALSE}
install.packages("devtools")
library(devtools)
install_github("ashiklom/pecan", subdir="modules/rtm")
# Defaults to branch 'master'. 
# For custom branches, add `ref = "branchname"`
```

This package relies on a modern (>= 2003) Fortran compiler, as determined by 
your R installation. On most Unix systems, this is standard, but R may specify 
a particular compiler version that you don't have, resulting in an installation 
error. To fix this, simply add the following to your system `~/R/Makevars` 
file.

```
FC = gfortran
```

# Overview of features

## Simulating reflectance

Available radiative transfer models are called by passing a vector of their 
parameters. Similar models with different versions (e.g. PROSPECT 4, 5, 5B) 
also have a "version" argument. These models return a matrix of reflectance, 
transmittance, and/or absorption spectra (depending on the model) at 1 nm 
resolution over the wavelength range 400 to 2500 nm.

The PROSPECT family of models returns the reflectance (column 1) and 
transmittance (column 2) for an individual leaf as a function of 4 to 6 
parameters (depending on the version).

```{r}
library(PEcAnRTM)
wl <- 400:2500
params <- c("N"=1.4, "Cab"=40, "Car"=15,
            "Cbrown"=0.5, "Cw"=0.002, "Cm"=0.004)

p4 <- prospect(params[c(-3,-4)], version=4)
p5 <- prospect(params[-4], version=5)
p5b <- prospect(params, version="5B")

plot(wl, p4[,1], type='l', xlab="Wavelength (nm)", ylab="Value", ylim=c(0,1))
lines(wl, 1-p4[,2], col=2)
lines(wl, p5[,1], lty=2, col=1)
lines(wl, 1-p5[,2], lty=2, col=2)
lines(wl, p5b[,1], lty=3, col=1)
lines(wl, 1-p5b[,2], lty=3, col=2)
legend("topright", c("Reflectance", "Transmittance"), col=c(1,2), lty=1)
legend("top", c("4", "5", "5B"), lty = c(1,2,3))
```

The SAIL family of models returns the bidirectional (1), hemispherical 
directional (2), directional hemispherical (3), and bidirectional hemispheical 
(4) reflectance factors for a canopy with a given set of approximately 20 
parameters. It is often coupled to the PROSPECT model as PRO4SAIL. 

```{r}
sail.params <- defparam("pro4sail")
print(sail.params)
p4s <- pro4sail(sail.params)
matplot(x = wl, y = p4s, type='l', xlab="Wavelength (nm)", ylab="Reflectance")
legend("topright", as.character(1:4), col=1:4, lty=1:4)
```

The above example illustrates the use of `defparam` to get the default 
parameters for a particular model. Similarly, `model.list` is a `data.table` 
containing all currently available models.

```{r}
print(model.list)
```


## Inversion

A novel feature of this package is the ability to perform a Bayesian inversion 
of a Radiative Transfer Model. Here are several advantages of the Bayesian 
approach:

* **Parameter uncertainty:** The output of a Bayesian analysis is a full joint 
  probability distribution of the model parameters, which includes a robust 
  estimate of their uncertainy and covariance between parameters.
* **Prior knowledge:** If previous, independent estimates of parameters are 
  available, these can be used to inform the model.
* **Partitioning variability:** Random effects models provide a powerful 
  framework for understanding the sources of variability and uncertainty in a 
  data set.

An inversion can be performed as follows:

1. Prepare your data. Observed spectra must be in a matrix with each row 
   corresponding to a wavelength and each column to an observation. An example 
   spectral matrix for _Acer rubrum_ leaves (`testspec_ACRU`) is provided in 
   `data(testspec)`.

```{r}
data(testspec)
observed <- testspec_ACRU[,1:3]
matplot(wl, observed, 
        xlab="Wavelength", ylab="Reflectance", type='l', lty=1)
```

2. Specify the model. The model takes the form of an R function with the 
   sampled parameters as input and a vector of equal length to the observations 
   as output.  A simple case is presented below; however, any arbitrary R 
   function will work.  Note that the constant `"Cbrown" = 0` is embedded in 
   the model.

```{r}
model <- function(params){
    in.params <- c(params["N"], params["Cab"], params["Car"],
                   "Cbrown"=0, params["Cw"], params["Cm"])
    out <- prospect(in.params, "5B")[,1]
    return(out)
}
```

3. Specify parameter priors. The prior is a single R function that takes the 
   parameter vector as an input and returns a single value: the log probability 
   density. Here, we provide an example where all the priors are lognormal (N 
   is a special case where the lognormal is translated to reflect its strict 
   lower bound at 1).

```{r}
mu <- c("N" = -0.427, "Cab"=3.332, "Car"=1.851, "Cw"=-4.499, "Cm"=-5.082)
sigma <- c("N" = 0.375, "Cab"=0.564, "Car"=0.761, "Cw"=0.546, "Cm"=0.656)
prior <- function(params){
    params["N"] <- params["N"] - 1
    return(sum(dlnorm(params, mu, sigma, log=TRUE)))
}
```

4. Set the initial conditions. Here, we do this by sampling randomly from the 
   prior. `params.prospect5` is a shortcut vector containing the names of the 
   PROSPECT 5 parameters.

```{r}
inits <- rlnorm(5, mu, sigma)
names(inits) <- params.prospect5
inits["N"] <- inits["N"] + 1
```

5. Set the minimum values for the inversion. When proposed parameters are below 
   this threshold, they are automatically rejected. For most parameters, these 
   will be 0.

```{r}
p5min <- c("N"=1, "Cab"=0, "Car"=0, "Cw"=0, "Cm"=0)
```

6. Perform inversion. The `do.lsq` parameter determines whether or not to 
   perform an optimized Least-Squares minimization first (this may accelerate 
   convergence, but runs the risk of trapping the inversion in a local 
   minimum). The `quiet` parameter determines whether or not to print inversion 
   progress and other status messages.

```{r}
samples <- invert.custom(observed = observed,
                         inits = inits,
                         ngibbs = 10000,
                         prior = prior,
                         pm = p5min,
                         model = model,
                         do.lsq = FALSE,
                         quiet=TRUE)
```



```{r}
# Analyze output
par(mfrow=c(3,2))
for(i in 1:6){
    param <- colnames(p5inv)[i]
    plot(p5inv[,i], type='l', main=param)
    p5.burnin <- p5inv[-500:0,i]
    plot(density(p5.burnin), main=param)
    m <- mean(p5.burnin)
    s <- sd(p5.burnin)
    abline(v=m)
    abline(v=m-s, col=2, lty=2)
    abline(v=m+s, col=2, lty=2)
}

print(colMeans(p5inv[-500:0,]))
print(apply(p5inv[-500:0,], 2, sd))
```

### Random effects

A more complicated analysis involves individual random effects.

```{r}
p5rand <- matrix(0, 5, 3)
rownames(p5rand) <- names(p5.means)
p5re <- invert.fast.re(modname = "prospect_5b",
                       observed = testspec_ACRU[,1:3],
                       inits = p5.means,
                       cons = c("Cbrown"=0),
                       rand = p5rand,
                       pmu = p5mu,
                       psd = p5sigma,
                       plog = p5log,
                       minp = p5min,
                       ngibbs = 1000)

print(colMeans(p5re[-50:0,]))
print(apply(p5re[-50:0,], 2, sd))
```
